---
title: "Lab 9b: Control Structures and Functions"
author: Grace Lawson
date: 10/28/2020
output: html_document
---

# Control Structures
Control structures allow the control of the flow of execution in a series of expressions. 
Commonly used structures:
- "if" and "else" = testing a condition and acting on it
- "for" = execute a loop a fixed number of times
- "while" = execute a loop while a condition is true

## Conditional Statements with "if-else"
```{r, eval = FALSE}
if(<condition>) {## do something}
  # continue with rest of code 
```

To execute an action when the condition is false, then use "else"
```{r, eval = FALSE}
if(<condition>) {
  ## do something
}
else {
  ## do something else
}
```

Can use a series of test by following initial "if" with other "else if"s
```{r, eval = FALSE}
if(<condition>) {
  ## do something
} else if(<condition>) {
  ## do something different
} else(<condition>) {
  ## do something different
}
```

### Example
```{r}
# Generate a uniform random number
x <- runif(1, 0, 10)
if(x > 3) {
  y <- 10
} else {
  y <- 0
}
```

## Important operators
### Comparison operators
- "==" (equal)
- "!=" (not equal)
- ">" (greater than)
- ">=" (greater than or equal to)
- "<" (less than)
- "<=" (less than or equal to)
### Logical operators
- "&" (and)
- "|" (or)
- "!" (not)

# "for" Loops
```{r}
for(i in 1:12) {
  print(i)
}
```

```{r}
x <- c("a", "b", "c", "d", "e")
for(i in 1:5) {
  print(x[i])
}
```

seq_along(x) makes an integer sequence based on length of the object
```{r}
for(i in seq_along(x)) {
  print(x[i])
}
```

```{r}
for(letter in x) {
  print(letter)
}
```

For one-line loops the curly braces aren't necessary:
```{r}
for(i in 1:3) print(x[i])
```

## Nested "for" loops
```{r}
x <- matrix(1:9, 3, 3)
for(i in seq_len(nrow(x))) {
  for(j in seq_len(ncol(x))) {
    print(x[i,j])
  }
}
```

Use nested loops for multidimensional or hierarchical data structures like lists or matrices. Try not to nest beyond 2 or 3 levels - instead, use functions.

## "while" loops 
"while" loops begin by testing a condition. If the condition is true, then they execute the loop. Once loop body is executed the condition is tested again, and this is repeated until the condition is false. 
```{r}
count <- 0
while(count < 10) {
  print(count)
  count <- count + 2
}
```

Be careful of accidentally creating an infinte loop

```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
  coin <- rbinom(1, 2, 0.5)
  
  if(coin == 1) {
    z <- z + 1
  } else {
    z <- z - 1
  }
}
print(z)
```

# Functions
Used to encapsulate a sequence of expressions that need to be executed numerous times

- Functions can be passed as arguments to other functions - useful for apply functions
- Functions can be nested 

Functions are defined using function() directive and are stored as objects 

## Example structure
```{r}
f <- function() {
  ## Empty function
}
# Execute function
f()
```

```{r}
f <- function() {
  cat("Hi!\n")
}
f()
```

Use function arguments to specify different options
```{r}
f <- function(num) {
  for(i in seq_len(num)) {
    cat("Hi!\n") 
  }
}
f(5)
```

```{r}
f <- function(num) {
  hello <- "Hello, world!\n"
  for(i in seq_len(num)) {
    cat(hello)
  }
  chars <- nchar(hello) * num
  chars
}
meaningoflife <- f(3)
print(meaningoflife)
```

The return value of the function is the last expression evaluated - chars variable is the last expression evaluated, so this is the return value.

User must specify value of the argument num - otherwise there will be an error.
We can modify this by setting a default value for num. 

```{r}
f <- function(num = 3) {
  hello <- "Hello, world!\n"
  for(i in seq_len(num)) {
    cat(hello)
  }
  chars <- nchar(hello) * num
  chars
}
f()
```
