---
title: "Lab_9"
author: Grace Lawson
date: 10.26.2020
output: html_document
---

```{r}
library(stringr)
library(tidyverse)
```

# Matching patterns with regular expressions
## Basic matches
Simplest patterns match exact strings
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```
"." matches any character except a new line
```{r}
str_view(x, ".a.")
```
- this only shows the first match in a string

To find an actual period, use a double backslash (ex - period = "\\.")
```{r}
dot <- "\\."
writeLines(dot)
```

```{r}
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```

To find an actual backslash, use four backslashes (ex \ = \\\\)
```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

### Exercises
1. Explain why each of these strings don't match a "\":
"\" - this is an escape for special behavior, not a regular expression for \
"\\" - this escapes the special behavior of \, but does not then recognize the character \
"\\\" - this escapes the special behavior of \, then has the string \, but doesn't escape the third backslash's special behavior

## Anchors
Regular expressions match any part of a string - can use anchors to direct it to match start or end of the string
```{r}
# Match start of the string with ^
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
# Match end of the string with $
str_view(x, "a$")
```

To force regular expression to match a complete string use both ^ and $
```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

### Exercises
1. How would you match the string "$^$"?
- str_view(x, "^$^$$")

## Character classes and alternatives
```{r}
# Look for a literal character that normally has a special regex meaning
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```
```{r}
str_view(x, ".[*]c")
```
```{r}
str_view(x, "a[ ]")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

### Exercises
1. Regex that:
- Start with a vowel = "^[aeiou]."
- Only contain consonants = "[aeiou]", negate = TRUE
- End with -ed but not -eed = "[^e]ed$"
- End with -ing or -ise = "i(ng|se)$"

## Repetition
Need to control how many times a pattern matches:
- ? = 0 or 1
- + = 1+
- * = 0+

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```
```{r}
str_view(x, "CC+")
```
```{r}
str_view(x, "C[LX]+")
```

Can also specify number of matches exactly:
- {n} = exactly n
- {n,} = n or more
- {,m} = at most, m
- {n,m} = between n and m

```{r}
str_view(x, "C{2}")
```
```{r}
str_view(x, "C{2,}")
```

To make these searches match the shortest string possible instead of the longest put a "?" after 
```{r}
str_view(x, "C{2,3}?")
```
```{r}
str_view(x, "C[LX]+?")
```

### Exercises
1. Describe the equivalents of ?,+,* in {m,n} form
- ? = {0,1}
- + = {1,}
- * = {0,}

## Grouping and backreferences
Parantheses create a numbered capturing group (number 1,2 etc). 
- Capturing group = stores part of string matched by the part of the regular expression inside parantheses. 
- Refer to the same text as previously matched using backreferences (\1, \2, etc)
```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

### Exercises
1. Describe what these expressions will match:
- (.)\1\1
    a character repeated 3 times in a row
- "(.)(.)\\2\\1"
    2 characters repeated in the opposite order
- (..)\1
    2 characters repeated
- "(.).\\1.\\1"
    character 1 followed by any character, then character 1, then any character, then character 1
- "(.)(.)(.).*\\3\\2\\1"
    3 characters, then some number of characters, then the former 3 in reverse order
    
# Tools
## Detect matches
To determine if character vector matches a pattern use str_detect() 
```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
# How many common words start with "t"?
sum(str_detect(words, "^t"))
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

When there are more complex conditions, combine calls with logical operators instead of creating a single regular expression

Example - find words that not don't have any vowels
```{r}
# Find all words containing at least one vowel, then negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

```{r}
# Find all words consisting only of consonants 
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

Select elements that match a pattern - can use logical subsetting or str_subset()
```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

For data frames - use filter
```{r}
df <- tibble(
    word = words,
    i = seq_along(word)
)
df %>% 
    filter(str_detect(word, "x$"))
```

str_count() tells how many matches there are in a string
```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

Use str_count() with mutate():
```{r}
df %>% 
    mutate(
        vowels = str_count(word, "[aeiou]"),
        consonants = str_count(word, "[^aeiou]")
    )
```

### Exercises
1. Find all words that start or end with x
Single regular expression
```{r}
words[str_detect(words, "^x|x$")]
```
Combination of str_detect() calls
```{r}
start_w_x <- str_detect(words, "^x")
end_w_x <- str_detect(words, "x$")
words[start_w_x | end_w_x]
```

## Extract matches
To extract the text of a match str_extract() 
Example - 
```{r}
length(sentences)
```
```{r}
head(sentences)
```

Find all sentences that contain a color & extract the color 
```{r}
colors <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match <- str_c(colors, collapse = " | ")
color_match
```

```{r}
has_color <- str_subset(sentences, color_match)
matches <- str_extract(has_color, color_match)
head(matches)
```

* str_extract() only extracts the first match
```{r}
more <- sentences[str_count(sentences, color_match) > 1]
str_view_all(more, color_match)
```
```{r}
str_extract(more, color_match)
```

* To get all matches, use str_extract_all()
```{r}
str_extract_all(more, color_match)
```

Use simplify = TRUE to return a matrix with short matches expanded to the same length as the longest

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

### Exercises
1. Modify regex to fix "flickered" match
```{r}
color_match_2 <- str_c("\\b(", str_c(colors, collapse = " | "), ")\\b")
color_match_2
```

## Grouped matches
Example - extract nouns from sentences
```{r}
noun <- "(a | the) ([^ ]+)"

has_noun <- sentences %>% 
    str_subset(noun) %>% 
    head(10)
has_noun %>% 
    str_extract(noun)
```

str_match gives each individual component - returns a matrix with one column for complete match followed by one column for each group
```{r}
has_noun %>% 
    str_match(noun)
```

If data is in a tibble - use tidyr::extract() 
```{r}
tibble(sentence = sentences) %>% 
    tidyr::extract(
        sentence, c("article", "noun"), "(a|the) ([^ ]+)",
        remove = FALSE
    )
```

### Exercises
1. Find words that come after a number - pull out number and word
```{r}
number_word <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, number_word)] %>% 
    str_extract(number_word)
```

## Replacing matches
Can replace matches with new strings 
```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```
```{r}
str_replace_all(x, "[aeiou]", "-")

```

Perform multiple replacements by supplying a named vector
```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

Use backreferences to insert components of the match
Example - flip the order of the second and third words
```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

### Exercise
1. Replace all forward slashes in a string with backslashes
```{r}
str_replace_all("a/b/c", "/", "\\\\")
```

## Splitting
Use str_split() to split a string into pieces
```{r}
sentences %>% 
    head(5) %>% 
    str_split(" ")
```

When working with a length-1 vector - just extract first element of list
```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

* Use simplify = TRUE to return a matrix
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

* Request a max number of pieces
```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

Split up by character, line, sentence, word using boundary()s
```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```
```{r}
str_split(x, " ")[[1]]
```
```{r}
str_split(x, boundary("word"))[[1]]
```

### Exercises
1. Split up a string into individual components 
```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```
